---
title: "US Metro Matching Tool Dashboard"
output: 
  flexdashboard::flex_dashboard:
    source: embed
    smart: false
runtime: shiny
---



```{r global, include=FALSE}

# data and libraries for matching tool

library( MatchIt )
library( here )
library( dplyr )
library( geojsonio )
library( sp )
library( pander )
library(here)


# data and libraries for shiny widgets

library( flexdashboard )
library( tidyverse )
library( shiny )
library( knitr )
library( rsconnect )
library( pander )
library( DT )

```


```{r, include=FALSE}


# Load the New Market Tax Credit raw data

df.raw <- read.csv(here("data/data-dashboard/data-nmtc-2019.csv"))

demo.data <- read.csv(here("data/data-dashboard/test-data.csv" ))

# Load the merged df - aggregated treatment data and census data

df.merged.treatment <- read.csv( here( "data/data-rodeo/rodeod-treatment-data.csv" ))

# Load the matched df - New Orleans matchit object

df3 <- read.csv(here("data/data-dashboard/new-orleans-matchit-object.csv"))
matching.results <- readRDS(here("data/data-dashboard/matching-results.rds"))

```



Demo - Step 1 (Download)
=====================================  


Inputs {.sidebar}
-------------------------------------

```{r}

h3( HTML("<b>Application Demo</b>") )

h4( "This application was designed to be used as a program evalution tool that identifies control group counterfactuals for a treatment program, using the genetic matching method in R. This dashboard will take you through the tool, step by step, using the New Market Tax Credit program as an example." )

h4( HTML("Follow the instructions for the following tabs: <br><hr><br>") )

h5( HTML( "<b>Tab 01</b>" ) )

h5(  "Download the raw New Market Tax Credit data as a .csv, so you can follow along. Make sure to use the dropdown menu to preview all to download the entire dataset." )

h5( HTML("<br><hr><br>") )

h5( HTML( "<b>Tab 02</b>" ) )

h5(  "Upload the .csv you just downloaded, and aggregate it to the census tract level." )

h5( HTML("<br><hr><br>") )

h5( HTML( "<b>Tab 03</b>" ) )

h5(  "Merge the data aggreagted on Tab 02 with census data from the Bureau of Census." )

h5( HTML("<br><hr><br>") )

h5( HTML( "<b>Tab 04</b>" ) )

h5(  "Run the matching function on the newly created dataframe." )

h5( HTML("<br><hr><br>") )

h5( HTML( "<b>Tab 05</b>" ) )

h5(  "Map the matches." )

h5( HTML("<br><hr><br>") )

#helpText("Additional information...")

```


Outputs
-------------------------------------

### 

```{r}

# renderDataTable({
#   
#   datatable( demo.data, 
#             filter = "bottom", 
#             rownames = FALSE, 
#             fillContainer = TRUE, 
#             style = "bootstrap",
#             class = "table-condensed table-striped",
#             extensions = "Buttons", 
#             options = list( dom = "Bfrtip", 
#                             buttons = "csv" ))
#   
#   
# 
#   })



renderDataTable({
  
  datatable( data = demo.data
                , extensions = 'Buttons'
                , options = list( 
                  dom = "Blfrtip"
                  , buttons = 
                    list("copy", list(
                      extend = "collection"
                      , buttons = c("csv", "excel", "pdf")
                      , text = "Download"
                    ) ) # end of buttons customization
                   
                   # customize the length menu
                  , lengthMenu = list( c(10, 20, -1) # declare values
                                       , c(10, 20, "All") # declare titles
                  ) # end of lengthMenu customization
                  , pageLength = 10
                   
                   
                ) # end of options
               
     ) # end of datatables
   

  
})

```



Demo - Step 2 (Upload and Aggregate) {data-orientation=rows}
===============================================================  


Inputs {.sidebar}
-------------------------------------

```{r}

fileInput( "demodata2", label = h3("Upload Test Data"),
           accept = c("text/csv",".csv") )

helpText("Instructions")

h5( HTML( "Upload the .csv you just downloaded on Tab 1, and aggregate it to the census tract level." ) )

```


Output Row 1
-------------------------------------

### 

```{r, eval=T}

# when input file changes
# then executes code 

df <- eventReactive( input$demodata2, {
        read.csv( input$demodata2$datapath )
      })


h3( "Raw Data from the File Upload" )

renderTable({

  # if (is.null( input$demodata2 ) )
  # return(NULL)
  
  dat <- df()
  head( dat )

})


```

Output Row 2
-------------------------------------

### 

```{r, eval=T}

h3( "Projects Aggregated by Census Tract" )

renderTable({
  
  dat <- df()
  
  # Preview aggregated treatment data

  treat.data <- 
    dat %>% 
    group_by( City, tractid ) %>% 
    summarize( num.of.projects=n(), tot.proj.spending=sum( ProjectCost, na.rm=TRUE ) ) %>% 
    arrange( City )

  td2 <<- treat.data
  
  head( treat.data, 15 )
  
})
```






Demo - Step 3 (Merge)
=====================================  


Inputs {.sidebar}
-------------------------------------

```{r}

# Action button (merge NMTC data with Census data)

actionButton( "merge", label = "Merge")


merge_button <- eventReactive( input$merge, {
        read.csv( input$demodata2$datapath )
      })

helpText("Instructions")

h5( HTML( "Click the merge button to merge the data aggregated on Tab 2 with census data from the Bureau of Census." ) )

  
```


Output Row 1
-------------------------------------

### 

```{r, eval=T}

 

h3( "Projects Aggregated by Census Tract" )

renderTable({
           
  dat <- merge_button()
  
  # Preview aggregated treatment data

  treat.data <- 
    dat %>% 
    group_by( City, tractid ) %>% 
    summarize( num.of.projects=n(), tot.proj.spending=sum( ProjectCost, na.rm=TRUE ) ) %>% 
    arrange( City )
  
  head( treat.data, 15 )
  
  
})


```



Output Row 2
-------------------------------------

### 

```{r, eval=T}


h3( "Aggregated Treatment data merged with Census data " )

renderTable({
           
  dat <- merge_button()
  
  # Preview aggregated treatment data
treat.data <-
    dat %>%
    group_by( City, tractid ) %>%
    summarize( num.of.projects=n(), tot.proj.spending=sum( ProjectCost, na.rm=TRUE ) ) %>%
    arrange( City )

  merged.data <- 
    # REPEATED CODE
    dat %>% 
    group_by( City, tractid ) %>% 
    summarize( num.of.projects=n(), tot.proj.spending=sum( ProjectCost, na.rm=TRUE ) ) %>% 
    arrange( City ) # %>% 
    # NEW CODE  
     #left_join( treat.data, df.merged.treatment, by.x = "tractid", by.y = "id2", all.x = TRUE )
     
    merged.data <-    merge(treat.data, df.merged.treatment, by.x = "tractid", by.y = "id2", all.x = TRUE)
      
    
  head( merged.data, 15 )
  
})

```




Demo - Step 4 (MatchIt)
=====================================  


Inputs {.sidebar}
-------------------------------------

```{r}
# Action button (Run MatchIt function)

actionButton( "get.matchit", label = "Run" )


get_matchit_results <- eventReactive( input$get.matchit, 
                { matching.results })

helpText("Instructions")

h5( HTML( "Click the run button to run the matching function on the newly created dataframe, and preview the statistical data for matches in New Orleans." ) )
  
```


Output Row 1
-------------------------------------

### 

```{r, eval=T}

h3( "Matching Model Results" )

renderPrint({
  
  matchit.results <- get_matchit_results()
  
  summary( matchit.results )
  
})

```


Output Row 2
-------------------------------------

### 

```{r, eval=T}

## Preview matched data of New Orleans

#h3( "New Orleans matched data" )

#renderTable({
  
#  dat <- df()
  
#  # Preview merged data - census data merged with aggregated treatment data

  
#})

h3( "Interpretation" )

h5(  "How to interpret these results?" )
  
h4( HTML("<br><hr><br>") )

h4( HTML( "<b>Call:</b>" ) )


# What do these arguments do? 
# ", , reestimate = TRUE, 
#    replace = FALSE, caliper = 0.25

h5( HTML('<b>method = "genetic":</b> This argument is calling the matching method called “genetic” matching, which is an algorithm that iteratively checks propensity scores. This matching method provides more balance than the other methods that could have been used.') )

h5( HTML('<b>discard = "both":</b> The discard argument specifies whether to discard units that are clear outliers, in this case we chose “both” so that both treated and control outliers would be discarded before the matching process begins.') )


h4( HTML("<br><hr><br>") )

h4( HTML( "<b>Summary of balance for all data:</b>" ) )

h5(  "First, we assess balance on the unmatched data, by looking at the distance row, as opposed to variable rows. Values of standardized mean differences and eCDF statistics close to zero indicate good balance, so we can see here that there is a good balance." )

h4( HTML("<br><hr><br>") )

h4( HTML( "<b>Summary of balance for matched data:</b>" ) )

h5(  "Next, we assess the balance of the matched data, again by the distance row of the table. Again, we can see that since these numbers are close to zero there is a good balance." )

h4( HTML("<br><hr><br>") )

h4( HTML( "<b>Percent Balance Improvement:</b>" ) )

h5(  "Finally, we assess the percent balance improvement between the matched data and the entire dataset. Values between 0 and 100 indicate that balance improved after matching, and values less than 0 indicate that balance got worse after matching. In this case, the mean difference improved by 98.3%, the eQQ median improved by 93.4%, and so on." )

h4( HTML("<br><hr><br>") )

h4( HTML( "<b>Sample sizes:</b>" ) )

h5(  "The sample size table shows the final matched groups. In total, there were 338 Control groups and 54 Treated groups. Of the 338 Control groups, 32 were matched, 297 went unmatched, and 9 were discarded. Of the 54 Treated groups, 32 were matched, 21 went unmatched (unable to finding appropriate Control group counterfactuals) and 1 was discarded (as an outlier)." )

```




Demo - Step 5 (Maps)
=====================================  


Inputs {.sidebar}
-------------------------------------
```{r}

helpText("Instructions")

h5( HTML( "Use the radio button to preview maps of matches found for 4 metropolitan cities." ) )

# Radio button to choose city

radioButtons( "demo.image", 
              label = h3("Cities"),
              choices = list( "Phoenix, AZ" = here("code/photos/phoenix.png") , 
                              "New Orleans, LA" = here("code/photos/new-orleans.png") , 
                              "Reno, NV" = here("code/photos/reno.png") , 
                              "New York, NY" = here("code/photos/new-york.png" )), 
              selected = here("code/photos/new-orleans.png")  
)



```




Outputs
-------------------------------------

### Neighborhood Matches by Census Data 


```{r, eval=F}
# DEVELOPMENT CODE
# TEST WIDGET INPUTS 
renderPrint({ input$demo.image  })  

get.image <- reactive({ input$demo.image })
renderPrint({ get.image() })
```




```{r, eval=F}
# static version
renderImage({
    list( src = "new-orleans.png" )
  }, deleteFile=FALSE )
```

```{r}
# dynamic version linked to widget

get.image <- reactive({ input$demo.image })

renderImage({
    list( src = get.image() )
  }, deleteFile=FALSE )
```


```{r, eval=F}
# EXAMPLE WHEN IMAGES IN THEIR OWN FOLDER

# Use an existing image, and don't delete the image after sending it
# NOTE: Image is in the subdirectory called 'images/'
# when input$n is 1 filename is ./images/image1.jpeg

renderImage({
  
    # When input$n is 1, 
    filename <- normalizePath(file.path('./images',
                              paste('image', input$n, '.jpeg', sep='')))

    # Return a list containing the filename
    list(src = filename)
    
}, deleteFile = FALSE)

```
Row 
-------------------------------------

### About this Dashboard

These dorling examples show genetic twins for treated census tracts that received New Markets Tax Credits (NMTC) between 2010-2017, and untreated census tracts that did not receive NMTC funding.

Cases are matched to identify twin untreated tracts in 2010 utilizing the census variables that measure:  

- Household Income 

- Percent of Persons with a College Degree 

- Percent of Persons who are White

Treated or untreated tracts by:

Matched or unmatched tracts 

The maps have matched data based on coloring dependent on groups:

- Blue - Control group matched

- Red - Treated group matched

- Dark Red - Treated group unmatched

- Grey - Control group unmatched 

If there is no counterfactual found, the map will produce a “No Valid Counterfactual Found.” 

About
===================================== 



Row 
-------------------------------------

### About this Dashboard

This dashboard was created by Melia Petersen and Jaesa Rogers, masters students in the Program Evaluation and Data Analytics program at the Watt's College of Public Service and Community Solutions at Arizona State University. 

This dashboard that has been coded in R for nonprofit workers to be able to compare the effects of programs on neighborhoods that have received the treatment versus neighborhoods that have not. This dashboard uses the New Markets Tax Credits as an example for the treatment variable and use all other census tracts in a metropolitan area that have not received treatment and code them as a control group. The treatment group is comprised of census tracts that have received either funding dollars, a program that has been implemented, or a policy applied to it, and the control group are the census tracts that have not had these programs applied to them. The purpose of this is to be able to analyze the impact (positive, negative or neutral) that a program has had on a given areas. 

Using the MatchIT package, this dashboard runs a genetic based matching on the census tracts, finding “twins” to show the impact that a program has had by comparing it to a census tract that is statistically the same. To visualize this information, dorling shapefiles have been created and show the treated, untreated, matched and unmatched tracts, demonstrating all of the data that is created through this process.


### Dashboard Authors

Melia Petersen, Melia.Petersen@asu.edu, Phoenix, AZ

Jaesa Rogers, Jaesa.Rogers@asu.edu, Phoenix, AZ

Client: Dr. Jesse Lecy, jlecy@asu.edu, Phoenix, AZ

Row
-------------------------------------

DATA DICTIONARY


```{r}
data.dictionary <- read.csv(here("data/data-dashboard/data-dictionary.csv"), stringsAsFactors=F )

data.dictionary$description <- stringi::stri_trans_general( data.dictionary$Description, "latin-ascii" )
data.dictionary %>%
  select( Column, Description ) %>%
  pander( )
```

