---
title: "US Metro Matching Tool Dashboard"
output: 
  flexdashboard::flex_dashboard:
    source: embed
    smart: false
runtime: shiny
---


```{r global, include=FALSE}

# data and libraries for matching tool

library( MatchIt )
library( here )
library( dplyr )
library( geojsonio )
library( sp )
library( pander )


# data and libraries for shiny widgets

library( flexdashboard )
library( tidyverse )
library( shiny )
library( knitr )
library( rsconnect )
library(pander)
library( DT )

```


```{r, include=FALSE}


# Load the New Market Tax Credit raw data

df.raw <- read.csv("LTDB_Std_2000_fullcount.csv")



# Load the New Market Tax Credit treatment data

df.treatment <- read.csv("rodeod-treatment-data.csv")

```



Demo - Step 1 (Preview)
=====================================  


Inputs {.sidebar}
-------------------------------------

```{r}

# Run buttons (Run Default - Preview and aggregate NMTC data)

actionButton("run.raw", label = "Show Raw NMTC data")

actionButton("run.rodeo", label = "Show Aggregated NMTC data")

```


Outputs
-------------------------------------

### 

```{r}


if (interactive()) {
 df.nmtc.raw <- eventReactive(input$run.raw, {
      head(df.raw)
    })
    
}


if (interactive()) {
 df.nmtc.rodeo <- eventReactive(input$run.rodeo, {
      head(df.treatment)
    })

  }

```


```{r}

## STEP 01 of Metro matching tool

# Load a catalog of URLs of dorling cartograms for each CBSA metro area in the US. 


load_crosswalk <- function( )
{
   dorlings.url <- "https://raw.githubusercontent.com/meliapetersen/matching-project/master/data/cbsa-names-and-shapefiles.csv"
   dorlings.catalog <- read.csv( dorlings.url )
   return( dorlings.catalog )
}

dorlings.catalog <- load_crosswalk()

# Load a dorling cartogram for one metro area: 


get_dorling <- function( cbsa, dorlings.catalog=NULL, mapit=FALSE )
{

   if( is.null(dorlings.catalog) ){ dorlings.catalog <- load_crosswalk() }


   file.name <- dorlings.catalog$shapefile[ dorlings.catalog$cbsa == cbsa ]
 

   base.url <- "https://raw.githubusercontent.com/DS4PS/usa-dorling-shapefiles/master/maps/metros-dorling/"

   full.url <- paste0( base.url, file.name )

   map <- geojsonio::geojson_read( x=full.url,  what="sp" )

   if( mapit ){ plot( map ) }

   return( map )
   
}


## STEP 02 of Metro matching tool



# Add the treatment data to the dorling spatial object: 

add_treatment <- function( df.treatment, dorling.object )
{

  # merge leftjoin dorling object and treatment data
  dorling.object <- merge( dorling.object, df.treatment, by.x="GEOID", by.y="id2", all.x=TRUE )


  # create treatment variable 
  treatment <- dorling.object$treatment 
  
  # omit NAs
  treatment[ is.na(treatment) ] <- 0
  
  #add treatment variable back into df
  dorling.object$treatment <- treatment
 

  return( dorling.object )
}



# Preview with New Orleans

  # call New Orleans dorling
  new.orleans.cbsa.code <- 35380
  new.orleans <- get_dorling( cbsa=new.orleans.cbsa.code )
  
  # merge New Orleans treatment data
  new.orleans <- add_treatment( df.treatment, dorling.object=new.orleans ) 
  


```



Demo - Step 2 (Merge)
=====================================  


Inputs {.sidebar}
-------------------------------------

```{r}
# Run button (Merge NMTC data and census data)


actionButton("merge", label = "Merge")



```


Outputs
-------------------------------------

### 


```{r}


# Preview existing treatment data

rmarkdown::paged_table(  head(df.treatment) )


# Preview New Orleans Merged data

rmarkdown::paged_table( head(new.orleans@data) )



```



Demo - Step 3 (Maps)
=====================================  


Inputs {.sidebar}
-------------------------------------
```{r}
# Radio button to choose city

radioButtons( "demo.image", 
              label = h3("Cities"),
              choices = list( "Phoenix, AZ" = "phoenix.png" , 
                              "Boulder, CO" = "boulder.png", 
                              "New Orleans, LA" = "new-orleans.png", 
                              "Reno, NV" = "reno.png", 
                              "New York, NY" = "new-york.png" ), 
              selected = "new-orleans.png" )

```




Outputs
-------------------------------------

### Neighborhood Matches by Census Data 


```{r, eval=F}
# DEVELOPMENT CODE
# TEST WIDGET INPUTS 
renderPrint({ input$demo.image  })  

get.image <- reactive({ input$demo.image })
renderPrint({ get.image() })
```




```{r, eval=F}
# static version
renderImage({
    list( src = "new-orleans.png" )
  }, deleteFile=FALSE )
```

```{r}
# dynamic version linked to widget

get.image <- reactive({ input$demo.image })

renderImage({
    list( src = get.image() )
  }, deleteFile=FALSE )
```


```{r, eval=F}
# EXAMPLE WHEN IMAGES IN THEIR OWN FOLDER

# Use an existing image, and don't delete the image after sending it
# NOTE: Image is in the subdirectory called 'images/'
# when input$n is 1 filename is ./images/image1.jpeg

renderImage({
  
    # When input$n is 1, 
    filename <- normalizePath(file.path('./images',
                              paste('image', input$n, '.jpeg', sep='')))

    # Return a list containing the filename
    list(src = filename)
    
}, deleteFile = FALSE)

```
Row 
-------------------------------------

### About this Dashboard

These dorling examples show genetic twins for treated census tracts that received New Markets Tax Credits (NMTC) between 2000-2017, and untreated census tracts that did not receive NMTC funding.

Cases are matched to identify twin untreated tracts in 2000 utilizing the census variables that measure:  

- Unemployment 

- College 

Treated or untreated tracts by:

Matched or unmatched tracts 

The maps have matched data based on coloring dependent on groups:

- Blue - Control group matched

- Red - Treated group matched

- Dark Red - Treated group unmatched

- Grey - Control group unmatched 

If there is no counterfactual found, the map will produce a “No Valid Counterfactual Found.” 
