---
title: "matching code"
author: "Melia Petersen"
date: "7/17/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message=F, warning=F )
```

```{r}
# need this installed to use
# genetic matching option in MatchIt
# install.packages("rgenoud")

# load package 
library( MatchIt )
library( here )
library( dplyr )
```

```{r}
#load LTDB data
d1 <- readRDS(here ( "data/data-rodeo/LTDB-2000.rds" ))
d2 <- readRDS(here ( "data/data-rodeo/LTDB-2010.rds" ))
md <- readRDS(here ( "data/data-rodeo/LTDB-META-DATA.rds" ))

#clean data 
d1 <- select( d1, - year )
d2 <- select( d2, - year )

d <- merge( d1, d2, by="tractid" )
d <- merge( d, md, by="tractid" )

#load NMTC data 
URL1 <- "https://raw.githubusercontent.com/DS4PS/cpp-528-spr-2020/master/labs/data/raw/NMTC/nmtc-sheet-01.csv"
nmtc <- read.csv( URL1, stringsAsFactors=F )

# remove anything not a number from the string
d$id2 <- gsub( "[^0-9]", "", d$tractid )

# fix IDs so they are match
d$id2 <- as.numeric( d$id2 )

# need to convert from currency to numeric
# current format: 
# head( nmtc$QLICI.Amount )
# [1] "$300,000.00 "   "$1,008,750.00 " "$977,000.00 "

# remove dollar sign and commas
nmtc$amount <- gsub( "[,$]", "", nmtc$QLICI.Amount )

# head(  nmtc$amount  )
# "300000.00 "  "1008750.00 " "977000.00 "

# convert characters to numeric 
nmtc$amount <- as.numeric( nmtc$amount ) %>% round(0)

# head(  nmtc$amount  )
# [1]  300000 1008750  977000

nmtc.dollars <- 
  nmtc %>% 
  filter( Origination.Year >= 2000 & Origination.Year <= 2010 ) %>%
  group_by( X2010.Census.Tract ) %>% 
  summarize( num.nmtc=n(), nmtc.total = sum( amount, na.rm=T ) )

d <- merge( d, nmtc.dollars, by.x="id2", by.y="X2010.Census.Tract", all.x=T )

# recode tracts that had no grants from NA to 0
d$num.nmtc[ is.na(d$num.nmtc) ] <- 0
d$nmtc.total[ is.na(d$nmtc.total) ] <- 0

#remove rural districts 
d <- filter( d, urban == "urban" )


```

```{r}
head (d)
```

### Create data frame for matching

```{r}
# create treatment variable
treatment <- ifelse(d$nmtc.total>0, 1, 0)

```


```{r}
# match nmtc as treatment (census tracts that recieved nmtc funding vs those that didn't)
# and seek balance on other SES characteristics 

df <- 
  d %>% 
  select( cbsa, treatment, rent00, hinc00, pnhwht12, pnhblk12, phisp12, 
              ppov12, punemp12,
              hinc12, hincb12, hincw12, 
              phs12, pcol12,
              pown12, mhmval12 ) %>% 
 filter( treatment %in% c( 1,0 ) ) %>% 
 mutate( Treated = as.numeric( treatment == 1 ), 
           Control =  as.numeric( treatment == 0 ))

# drop rows with missing values

nrow( df )
df2 <- na.omit( df )
nrow( df2 )


# "treated" group is white
# comparison set is minority communities 

head( df2)

```


### matchit function 

### Genetic Search Method

Better method, but more computationally intensive. 

We use a smaller sample here, but to run with the full sample it may require running it overnight or on the supercomputing clusters. 


**NOTE** the long output printed while the algorithm runs. Add code chunk option

results 

```{r, results="hide", cache=TRUE}
# test genetic search method
# is time-intensive so test with a sample

# set seeds if you want to reproduce the 
# exact result in the future
set.seed( seed=1234 )

# create sample set for top NMTC 

df3 <- df2  %>%  filter(cbsa == c(35380, 20260, 31100, 16980, 33340 ))
                 
                 
table( df3$treatment )


m3 <- matchit( mhmval12 ~ hinc12 + pcol12 + pnhwht12, 
               method="genetic", discard="both", reestimate=TRUE, 
               replace=TRUE, data = df3 ) 

m4 <- matchit( mhmval12 ~ hinc12 + pcol12 + pnhwht12, 
               method="genetic", discard="both", reestimate=TRUE, 
               replace=TRUE, caliper=.25, data = df3 ) 

m5 <- matchit( mhmval12 ~ hinc12 + pcol12 + pnhwht12, 
               method="genetic", discard="both", reestimate=TRUE, 
               replace=FALSE, caliper=.25, data = df3 ) 

```


Report resulting fits:

```{r}
summary( m3 )   # replacement true no calipers 
summary( m4 )   # replacement true with calipers 
summary( m5 )   # replacement false with calipers 
```



### Notes on Arguments 

```{r}
# NOTES
# two stages of dropping observations:
#
# FIRST STAGE:
# discard drops cases outside of common support before matching starts
# reestimate re-calculates distances after discards are made 
#
# SECOND STAGE:
# replace=TRUE is the default, which means each treated case
#   is matched with the closest comparison case, but those
#   can be repeated, so one untreated case can serve as the 
#   twin for several treated cases and is duplicated in the final dataset
#   if replace=FALSE then each comparison case is only used once
#
# after matches are made caliper determines which matches are kept;
# it measures standard deviations apart, 
# so 0.25 allows a max distance of 0.25 std devs for any covariate X
# and drops all cases where 1 or more covariates fall outside this distance


# NOTES: balance achieved but with heavy resampling -
# comparison group members reused on average almost three times
# (523/187) to provide balance 
# 
# m3 <- matchit( white ~ hinc12 + phs12 + mhmval12, 
#                method="genetic", discard="both", reestimate=TRUE, 
#                replace=TRUE, data = df3 ) 
# 
# Summary of balance for matched data:
#          Means Treated Means Control  SD Control Mean Diff   eQQ Med   eQQ Mean     eQQ Max
# distance        0.6589        0.6595      0.1988   -0.0006 9.370e-02     0.0917      0.1416
# hinc12      62014.9388    61848.5310  20112.9137  166.4078 9.103e+03  8332.4667  24342.0000
# phs12          41.6632       41.6850     13.9822   -0.0218 3.160e+00     3.1851      7.2500
# mhmval12   200500.9579   198128.3723 106198.3885 2372.5856 1.740e+04 21014.7715 353301.0000
# 
# Percent Balance Improvement:
#          Mean Diff. eQQ Med eQQ Mean   eQQ Max
# distance    99.7863 70.8770  67.5513   60.1371
# hinc12      99.2302 59.5853  61.1588   61.1640
# phs12       99.8260 75.3318  74.6637   70.6952
# mhmval12    86.7505 34.0909  17.1436 -244.3480
# 
# Sample sizes:
#           Control Treated
# All           469     531
# Matched       187     523
# Unmatched     281       0
# Discarded       1       8
# 
# 

# NOTES: add caliper, so we increase unmatched cases 
# by dropping those with poor fit after matching,
# but we still have replacement so we reuse comparison 
# cases on average about two times (361/163)
# 
# m4 <- matchit( white ~ hinc12 + phs12 + mhmval12, 
#                method="genetic", discard="both", reestimate=TRUE, 
#                replace=TRUE, caliper=.25, data = df3 ) 
# 
# Sample sizes:
#           Control Treated
# All           469     531
# Matched       163     361
# Unmatched     305     162
# Discarded       1       8


# CAUTION! If you change replace to FALSE but do not add caliper
# then you will almost certainly get an unbalanced dataset
# because you will try to match every treated case with its
# nearest neighbor, but the comparison cases closest to the 
# treated cases have already been taken, so the further you go down
# the line the worst the matches get

# NOTES: replace = FALSE means each comparison case can only be
# used one time, then caliper = 0.25 drops poor matches,
# so this returns the most restrictive sample - only cases
# that were closely matched but without allowing duplication 
# of comparison cases 
#
#
# m5 <- matchit( white ~ hinc12 + phs12 + mhmval12, 
#                method="genetic", discard="both", reestimate=TRUE, 
#                replace=FALSE, caliper=.25, data = df3 ) 
# 
# Sample sizes:
#           Control Treated
# All           469     531
# Matched       197     197
# Unmatched     271     326
# Discarded       1       8
```


Constructing the balanced sample: 

Examples here are fairly manual to show what information the matchit return object contains. 

```{r}
names( m3 )
```


```{r}
# match matrix is the length of the original treated groups
# IDs are rows of the original dataset

matches <- data.frame( treatment.group.id=row.names( m5$match.matrix), 
                       control.group.id=m5$match.matrix )

head( matches, 25 )

# compare quality of matches
# select id (row) of treated and id (row) of comparison,
# compare on model covariates:

# note the nice balance: 

df3[ c(2,886) , c("white","hinc12","phs12","mhmval12") ]


m5$nn

# all treated cases in dataset:
length( m5$match.matrix )

# treated cases that were matched: 
sum( ! is.na(m5$match.matrix) )
```


### Extract balanced dataset 

Keep only cases that were matched in previous steps 

```{r}
new.dat <- match.data( m5 )
dim( new.dat ) # 2 x number in matched treatment group
head( new.dat )
```








### Set to Eval=F



```{r, eval=F}
#MatchIt code 

m1 = matchit( df$school ~ df$income + df$occ_score + df$y_educ, method = "nearest", data = d )

#code interpretation from notes
matchit(program(y/n/) ~ ses + race + age(?), method = "nearest", data = data) 

```

```{r, eval=F}

summary(m1)  # print balance of groups 
plot( m1 )   # visual comparison 

new.dat <- match.data( m1 )

```




