---
title: "New Market Tax Credit Genetic Twins for Metropolitan US Cities"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    theme: readable
    df_print: paged
    highlight: tango
    css: clean.css
---

```{r setup, include=FALSE}
knitr::opts_chunk$set( message=F, warning=F, fig.width=12, fig.height=12 )
```


```{r}
# need this installed to use
# genetic matching option in MatchIt
# install.packages("rgenoud")

# load package 
library( MatchIt )
library( here )
library( dplyr )
library( geojsonio )
library( sp )
library( pander )
```


This code walks through how to make genetic twins for treated census tracts that received New Market Tax Credits (NMTC) between 2000-2017, utilizing dorling shape files that were created for this project and can be found here: https://github.com/DS4PS/usa-dorling-shapefiles/tree/master/maps/metros-dorling 

Cases are matched to identify twin untreated tracts in 2000 utilizing the census variables that measure:  
- Unemployment
- College 

We have created a data frame that combines census data and NMTC data, assigning the treatment variable to the NMTC tracts, and control to those that have not received funding under this program.


## STEP 01

Load a catalog of URLs of dorling cartograms for each CBSA metro area in the US. 

```{r}
load_crosswalk <- function( )
{
   dorlings.url <- "https://raw.githubusercontent.com/meliapetersen/matching-project/master/data/cbsa-names-and-shapefiles.csv"
   dorlings.catalog <- read.csv( dorlings.url )
   return( dorlings.catalog )
}


```


```{r}
dorlings.catalog <- load_crosswalk()
rmarkdown::paged_table( head( dorlings.catalog ) )
```

Load a dorling cartogram for one metro area: 

```{r}
get_dorling <- function( cbsa, dorlings.catalog=NULL, mapit=FALSE )
{

   if( is.null(dorlings.catalog) ){ dorlings.catalog <- load_crosswalk() }


   file.name <- dorlings.catalog$shapefile[ dorlings.catalog$cbsa == cbsa ]
 

   base.url <- "https://raw.githubusercontent.com/DS4PS/usa-dorling-shapefiles/master/maps/metros-dorling/"

   full.url <- paste0( base.url, file.name )

   map <- geojsonio::geojson_read( x=full.url,  what="sp" )

   if( mapit ){ plot( map ) }

   return( map )
   
}
```

Test the function with New Orleans data: 

```{r}
new.orleans.cbsa.code <- 10580
new.orleans <- get_dorling( cbsa=new.orleans.cbsa.code )
```


## STEP 02

Load the New Market Tax Credit treatment data: 

```{r}
df.treatment <- read.csv("rodeod-treatment-data.csv")
rmarkdown::paged_table(  head(df.treatment) )
```


Add the treatment data to the dorling spatial object: 

```{r}
add_treatment <- function( df.treatment, dorling.object )
{

  # merge leftjoin dorling object and treatment data
  dorling.object <- merge( dorling.object, df.treatment, by.x="GEOID", by.y="id2", all.x=TRUE )


  # create treatment variable 
  treatment <- dorling.object$treatment 
  
  # omit NAs
  treatment[ is.na(treatment) ] <- 0
  
  #add treatment variable back into df
  dorling.object$treatment <- treatment
 

  return( dorling.object )
}
```



Test the function:

```{r}
new.orleans <- add_treatment( df.treatment, dorling.object=new.orleans ) 

# unit test - make sure data merge worked
table( new.orleans$treatment, useNA="ifany" )

# appended data 
rmarkdown::paged_table( head(new.orleans@data) )
```


## STEP 03

Match the "treated" census tracts (those that received federal aid) with *equivalent* untreated tracts using the matchit package.


```{r}
match_tracts <- function( fo, dorling.object )
{
   # extract data frame from sp object
   # rename rows using tract IDs
   
   df <- as.data.frame( dorling.object )
   row.names(df) <- df$GEOID
   
   # convert string to formula object
   fo <- as.formula( fo )
   
   matchit.object <- NULL

   # find matches in dataframe based on Household income, college education, and whiteness 
   try( 
      
      invisible({ capture.output({ 
         
      matchit.object <- matchit( fo, 
                        method="genetic", discard="both", reestimate=TRUE,
                        replace=FALSE, caliper=.25, data = df )
      })  }),
      
      
      silent=TRUE )

   return( matchit.object )

}
```

Test the function: 

```{r}
# define the matching function formula 
fo <- as.formula( "treatment ~ COL00 + EMP00" )

new.orleans.matches <- 
   match_tracts( fo, new.orleans ) 

summary( new.orleans.matches )
```


## STEP 04

Assign tracts to groups to treated or untreated tracts by matched or unmatched tracts. 


Treated or untreated tracts by
Matched or unmatched tracts 

```{r}
assign_groups <-  function( matchit.object, dorling.object )
{

  if( is.null( matchit.object ) )
  { 
     status <- rep( "control.unmatched", nrow( dorling.object ) )
     status[ dorling.object$treatment == 1 ] <- "treatment.unmatched"
     dorling.object$status <- status
  }
   
  if( ! is.null( matchit.object ) )  
  {

     # extract matched pairs from matchit object
     # match matrix is the length of the original treated groups
     # IDs are rows of the original dataset
     
     matches <- data.frame( treatment.group.id=row.names( matchit.object$match.matrix), 
                            control.group.id=matchit.object$match.matrix )
   
     these.na <- is.na( matches$control.group.id )
   
     # generate four cases for tracts 
     id.all.treated <- matches$treatment.group.id 
     id.treated.matched <- matches$treatment.group.id[ ! these.na ]
     id.treated.unmatched <- matches$treatment.group.id[ these.na ]
     id.control.matched <- na.omit(matches$control.group.id)
     id.control.unmatched <- dorling.object$GEOID [ !( dorling.object$GEOID %in% id.all.treated | 
                                                    dorling.object$GEOID %in% id.control.matched ) ]
   
     ### Assign tract IDs to groups
     
     treated.matched <- na.omit(id.treated.matched) %>% as.numeric
     control.matched <- na.omit(id.control.matched) %>% as.numeric
     
     treated.unmatched <- na.omit(id.treated.unmatched)  %>% as.numeric
     control.unmatched <- na.omit(id.control.unmatched)  %>% as.numeric
     
     status <- rep( "control.unmatched", nrow( dorling.object ) )
     status[ dorling.object$GEOID %in% control.matched ] <- "control.matched"
     status[ dorling.object$GEOID %in% treated.matched ] <- "treatment.matched"
     status[ dorling.object$GEOID %in% treated.unmatched ] <- "treatment.unmatched"
     dorling.object$status <- status
      
  }
   
  return( dorling.object )

}
```

```{r}
new.orleans <- 
   assign_groups( matchit.object=new.orleans.matches, 
                  dorling.object=new.orleans )

table( new.orleans$status )
```

# STEP 05 - Map Matches 

Map the matched data based on coloring dependent on groups: 
Blue - Control group matched 
Red - Treated group matched
Dark Red - Treated group unmatched 
Grey - Control group unmatched 

If there is no counterfactual found, the map will produce a “No Valid Counterfactual Found.” 


```{r}
map_matches <- function( dorling.object, cbsa.name, zoom=TRUE, zoom.range=90 )
{

   
   if( sum( dorling.object$status == "treatment.matched" ) > 0 )
   {
      
   ### Create color vectors based on matches
   
   color.vector <- rep( "gray80", nrow( dorling.object ) )
   color.vector[ dorling.object$status == "control.matched" ] <- "blue"
   color.vector[ dorling.object$status == "treatment.matched" ] <- "darkred"
   color.vector[ dorling.object$status == "treatment.unmatched" ] <- adjustcolor( "darkred", alpha.f = 0.4 )
   
   border.color <- rep( "gray40", nrow(dorling.object) )
   border.color[ dorling.object$status == "treatment.unmatched" ] <- "darkred"
   
   
   # set zoom
   
   coords <- coordinates( dorling.object ) %>% as.data.frame()
   
   x.min.max <- range( coords$V1, na.rm=T )
   y.min.max <- range( coords$V2, na.rm=T )
   
   if( zoom )
   {
     tail <- ( 100 - zoom.range ) / 200   
     x.min.max <- quantile( coords$V1, c( 0+tail, 1-tail ) )
     y.min.max <- quantile( coords$V2, c( 0+tail, 1-tail ) )
   }
   
   par( mar=c(0,0,5,0) )
   
   plot( dorling.object, 
         col=color.vector, 
         border=border.color, 
         xlim=x.min.max,
         ylim=y.min.max,
         main=toupper(cbsa.name), 
         cex.main=1.5 )
   }
   
   
       if( sum(dorling.object$treatment, na.rm = TRUE) < 1) 
   {   
     cat( "<br><br><hr><br><br>" )
     cat( paste0( "<center><h2>","There are no treated tracts in ", toupper(cbsa.name), ".</h2></center>") )
     cat( "<br><br><hr><br><br>" )
     return( NULL )
   }
   
   if( sum( dorling.object$status == "treatment.matched" ) == 0 )
   {
      par( mar=c(0,0,5,0) )
      plot( dorling.object,
            col=ifelse( dorling.object$treatment == 1, "darkred", "gray80" ),
            main=paste0( toupper(cbsa.name), ": \nNO VALID COUNTERFACTUALS FOUND" ), 
            cex.main=1.5 )
            
   }
   
   
      
}

```

```{r}
map_matches( new.orleans, cbsa.name="New Orleans", zoom=TRUE, zoom.range=90 )
```

```{r}


```



# Combine Steps 

After completing the steps to creating the functions that will find matched twins for the treated groups, you will combine all steps to prepare to loop for all metro CBSAs.


```{r}
get_matches <- function( cbsa.code, cbsa.catalog, 
                         df.treatment, fo,
                         zoom.range=90, return.dorling=FALSE )
{
   
   cbsa.name <- cbsa.catalog$cbsaname[ cbsa.catalog$cbsa == cbsa.code ]
   dorling.object <- get_dorling( cbsa=cbsa.code, cbsa.catalog )
   dorling.object <- add_treatment( df.treatment, dorling.object )
   matched.tracts <- match_tracts( fo=fo, dorling.object )
   
   cat( "<br><br><hr><br><br>" )
   cat( paste0( "<center><h2>", toupper(cbsa.name), "</h2></center>") )
   
   if( ! is.null(matched.tracts) )
   { 
      cat( matched.tracts$nn %>% pander() ) 
   }

   
   dorling.object <- assign_groups( matched.tracts, dorling.object )
   map_matches( dorling.object, cbsa.name, zoom=TRUE, zoom.range )
   
   if( return.dorling ){ return( dorling.object ) }
   
}
```


```{r, eval=F}
fo <- as.formula( "treatment ~ unemp00 + col00" )

akron <- 
   get_matches( cbsa.code=10420, 
                cbsa.catalog=dorlings.catalog,
                df.treatment=df.treatment,
                fo=fo,
                return.dorling=TRUE )

table( akron$status, useNA="ifany" )
map_matches( akron, cbsa.name="AKRON, OH", zoom=FALSE )
```


## Create loop for all CBSAs 

```{r, results="asis"}

# dorlings.catalog <- load_crosswalk()

# df.treatment <- # load treatment data 

# matching formula 
fo <- as.formula( "treatment ~  unemp00 + col00" )
cbsa.code.list <- unique( dorlings.catalog$cbsa )

for( i in cbsa.code.list )
{
  get_matches( cbsa.code=i, 
               cbsa.catalog=dorlings.catalog, 
               df.treatment=df.treatment, 
               fo=fo )
}


```




```{r}
fo <- as.formula( "treatment ~ unemp00 + col00" )
cbsa.code.list <- unique( dorlings.catalog$cbsa )

for( i in cbsa.code.list [193:349] )
{
  get_matches( cbsa.code=i, 
               cbsa.catalog=dorlings.catalog, 
               df.treatment=df.treatment, 
               fo=fo )
}
```



```{r}

fo <- as.formula( "treatment ~ unemp00 + col00" )
cbsa.code.list <- unique( dorlings.catalog$cbsa )

for( i in cbsa.code.list [196:349] )
{
  get_matches( cbsa.code=i, 
               cbsa.catalog=dorlings.catalog, 
               df.treatment=df.treatment, 
               fo=fo )
}
```


```{r}

fo <- as.formula( "treatment ~ unemp00 + col00" )
cbsa.code.list <- unique( dorlings.catalog$cbsa )

for( i in cbsa.code.list [208:349] )
{
  get_matches( cbsa.code=i, 
               cbsa.catalog=dorlings.catalog, 
               df.treatment=df.treatment, 
               fo=fo )
}
```

```{r}

fo <- as.formula( "treatment ~ unemp00 + col00" )
cbsa.code.list <- unique( dorlings.catalog$cbsa )

for( i in cbsa.code.list [295:349] )
{
  get_matches( cbsa.code=i, 
               cbsa.catalog=dorlings.catalog, 
               df.treatment=df.treatment,
               fo=fo )
}
```



