---
title: "metro-matching-loop"
author: "Jaesa Rogers"
date: "9/14/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message=F, warning=F, fig.width=10 )
```


```{r}
# need this installed to use
# genetic matching option in MatchIt
# install.packages("rgenoud")

# load package 
library( MatchIt )
library( here )
library( dplyr )

library( geojsonio )
library( sp )
```


## STEP 01

Load a catalog of URLs of dorling cartograms for each CBSA metro area in the US. 

```{r}
load_crosswalk <- function( )
{
   dorlings.url <- "https://raw.githubusercontent.com/meliapetersen/matching-project/master/data/cbsa-names-and-shapefiles.csv"
   dorlings.catalog <- read.csv( dorlings.url )
   return( dorlings.catalog )
}


```


```{r}
dorlings.catalog <- load_crosswalk()
head( dorlings.catalog )
```

Load a dorling cartogram for one metro area: 

```{r}
get_dorling <- function( cbsa, dorlings.catalog=NULL, mapit=FALSE )
{

   if( is.null(dorlings.catalog) ){ dorlings.catalog <- load_crosswalk() }


   file.name <- dorlings.catalog$shapefile[ dorlings.catalog$cbsa == cbsa ]
 

   base.url <- "https://raw.githubusercontent.com/DS4PS/usa-dorling-shapefiles/master/maps/metros-dorling/"

   full.url <- paste0( base.url, file.name )

   map <- geojsonio::geojson_read( x=full.url,  what="sp" )

   if( mapit ){ plot( map ) }

   return( map )
   
}
```

Test the function with New Orleans data: 

```{r}
new.orleans.cbsa.code <- 10580
new.orleans <- get_dorling( cbsa=new.orleans.cbsa.code )
```


## STEP 02

Load the treatment data: 

```{r}
df.treatment <- read.csv("rodeod-treatment-data.csv")
head(df.treatment)
```


Add the treatment data to the dorling spatial object: 

```{r}
add_treatment <- function( df.treatment, dorling.object )
{

  # merge leftjoin
  dorling.object <- merge( dorling.object, df.treatment, by.x="GEOID", by.y="id2", all.x=TRUE )


  # dorling.object <- df2
  # x <- c( 1, NA, NA, 1, NA, NA )
  # 
  # omit NAs
  treatment <- dorling.object$treatment 
  treatment[ is.na(treatment) ] <- 0
  dorling.object$treatment <- treatment
 
  # row.names(df.rodeo.metro) <- df.rodeo.metro$id2    
  # df.rodeo.metro <- na.omit( df.rodeo.metro )

  return( dorling.object )
}
```



Test the function:

```{r}
new.orleans <- add_treatment( df.treatment, dorling.object=new.orleans ) 

# unit test - make sure data merge worked
table( new.orleans$treatment, useNA="ifany" )

# appended data 
head(new.orleans@data)
```


## STEP 03

Match the "treated" census tracts (those that received federal aid) with *equivalent* untreated tracts using the matchit package.


```{r}
match_tracts <- function( fo, dorling.object )
{
   # extract data frame from sp object
   # rename rows using tract IDs
   
   df <- as.data.frame( dorling.object )
   row.names(df) <- df$GEOID
   
   # convert string to formula object
   fo <- as.formula( fo )
   
   matchit.object <- NULL

   # find matches in dataframe based on Household income, college education, and whiteness 
   try( 
      invisible({ capture.output({ 
         
      matchit.object <- matchit( fo, 
                        method="genetic", discard="both", reestimate=TRUE,
                        replace=FALSE, caliper=.25, data = df )
      })  })
   )

   return( matchit.object )

}
```

Test the function: 

```{r}
# define the matching function formula 
fo <- as.formula( "treatment ~ own00" )

new.orleans.matches <- 
   match_tracts( fo, new.orleans ) 

summary( new.orleans.matches )
```


## STEP 04

### Assign tracts to groups

Treated or untreated tracts by
Matched or unmatched tracts 

```{r}
assign_groups <-  function( matchit.object, dorling.object )
{

  if( is.null( matchit.object ) )
  { 
     status <- rep( "control.unmatched", nrow( dorling.object ) )
     status[ dorling.object$treatment == 1 ] <- "treatment.unmatched"
     dorling.object$status <- status
  }
   
  if( ! is.null( matchit.object ) )  
  {

     # extract matched pairs from matchit object
     # match matrix is the length of the original treated groups
     # IDs are rows of the original dataset
     
     matches <- data.frame( treatment.group.id=row.names( matchit.object$match.matrix), 
                            control.group.id=matchit.object$match.matrix )
   
     these.na <- is.na( matches$control.group.id )
   
     # generate four cases for tracts 
     id.all.treated <- matches$treatment.group.id 
     id.treated.matched <- matches$treatment.group.id[ ! these.na ]
     id.treated.unmatched <- matches$treatment.group.id[ these.na ]
     id.control.matched <- na.omit(matches$control.group.id)
     id.control.unmatched <- dorling.object$GEOID [ !( dorling.object$GEOID %in% id.all.treated | 
                                                    dorling.object$GEOID %in% id.control.matched ) ]
   
     ### Assign tract IDs to groups
     
     treated.matched <- na.omit(id.treated.matched) %>% as.numeric
     control.matched <- na.omit(id.control.matched) %>% as.numeric
     
     treated.unmatched <- na.omit(id.treated.unmatched)  %>% as.numeric
     control.unmatched <- na.omit(id.control.unmatched)  %>% as.numeric
     
     status <- rep( "control.unmatched", nrow( dorling.object ) )
     status[ dorling.object$GEOID %in% control.matched ] <- "control.matched"
     status[ dorling.object$GEOID %in% treated.matched ] <- "treatment.matched"
     status[ dorling.object$GEOID %in% treated.unmatched ] <- "treatment.unmatched"
     dorling.object$status <- status
      
  }
   
  return( dorling.object )

}
```

```{r}
new.orleans <- 
   assign_groups( matchit.object=new.orleans.matches, 
                  dorling.object=new.orleans )

table( new.orleans$status )
```

# STEP 05 - Map Matches 

```{r}
map_matches <- function( dorling.object, cbsa.name, zoom=TRUE, zoom.range=90 )
{

   
   if( sum( dorling.object$status == "treatment.matched" ) > 0 )
   {
      
   ### Create color vectors based on matches
   
   color.vector <- rep( "gray80", nrow( dorling.object ) )
   color.vector[ dorling.object$status == "control.matched" ] <- "blue"
   color.vector[ dorling.object$status == "treatment.matched" ] <- "darkred"
   color.vector[ dorling.object$status == "treatment.unmatched" ] <- adjustcolor( "darkred", alpha.f = 0.4 )
   
   border.color <- rep( "gray40", nrow(dorling.object) )
   border.color[ dorling.object$status == "treatment.unmatched" ] <- "darkred"
   
   
   # set zoom
   
   coords <- coordinates( dorling.object ) %>% as.data.frame()
   
   x.min.max <- range( coords$V1, na.rm=T )
   y.min.max <- range( coords$V2, na.rm=T )
   
   if( zoom )
   {
     tail <- ( 100 - zoom.range ) / 200   
     x.min.max <- quantile( coords$V1, c( 0+tail, 1-tail ) )
     y.min.max <- quantile( coords$V2, c( 0+tail, 1-tail ) )
   }
   
   par( mar=c(0,0,5,0) )
   
   plot( dorling.object, 
         col=color.vector, 
         border=border.color, 
         xlim=x.min.max,
         ylim=y.min.max,
         main=toupper(cbsa.name), 
         cex.main=1.5 )
   }
   
   if( sum( dorling.object$status == "treatment.matched" ) == 0 )
   {
      par( mar=c(0,0,5,0) )
      plot( dorling.object,
            col=ifelse( dorling.object$treatment == 1, "darkred", "gray80" ),
            main=paste0( toupper(cbsa.name), ": \nNO VALID COUNTERFACTUALS FOUND" ), 
            cex.main=1.5 )
            
   }
   
}

```

```{r}
map_matches( new.orleans, cbsa.name="New Orleans", zoom=TRUE, zoom.range=90 )
```




# Combine Steps 

```{r}
get_matches <- function( cbsa.code, cbsa.catalog, 
                         df.treatment, fo,
                         zoom.range=90, return.dorling=FALSE )
{
   
   cbsa.name <- cbsa.catalog$cbsaname[ cbsa.catalog$cbsa == cbsa.code ]
   dorling.object <- get_dorling( cbsa=cbsa.code, cbsa.catalog )
   dorling.object <- add_treatment( df.treatment, dorling.object )
   matched.tracts <- match_tracts( fo=fo, dorling.object )
   if( ! is.null(matched.tracts) ){ print(summary( matched.tracts )) }
   dorling.object <- assign_groups( matched.tracts, dorling.object )
   map_matches( dorling.object, cbsa.name, zoom=TRUE, zoom.range )
   
   if( return.dorling ){ return( dorling.object ) }
   
}
```


```{r, eval=F}
fo <- as.formula( "treatment ~ own00" )

akron <- 
   get_matches( cbsa.code=10420, 
                cbsa.catalog=dorlings.catalog,
                df.treatment=df.treatment,
                fo=fo,
                return.dorling=TRUE )

table( akron$status, useNA="ifany" )
map_matches( akron, cbsa.name="AKRON, OH", zoom=FALSE )
```


## Create loop for all CBSAs 

```{r}

# dorlings.catalog <- load_crosswalk()

# df.treatment <- # load treatment data 

# matching formula 
fo <- as.formula( "treatment ~ own00" )
cbsa.code.list <- unique( dorlings.catalog$cbsa )

for( i in cbsa.code.list[1:5] )
{
  get_matches( cbsa.code=i, 
               cbsa.catalog=dorlings.catalog, 
               df.treatment=df.treatment, 
               fo=fo )
}


```




