---
title: "New-Orleans-matching"
author: "Jaesa Rogers"
date: "8/2/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message=F, warning=F )
```

# Create New Orelans matches

```{r}
# need this installed to use
# genetic matching option in MatchIt
# install.packages("rgenoud")

# load package 
library( MatchIt )
library( here )
library( dplyr )
```

```{r}
#load LTDB data
d1 <- readRDS(here ( "data/data-rodeo/LTDB-2000.rds" ))
d2 <- readRDS(here ( "data/data-rodeo/LTDB-2010.rds" ))
md <- readRDS(here ( "data/data-rodeo/LTDB-META-DATA.rds" ))

#clean data 
d1 <- select( d1, - year )
d2 <- select( d2, - year )

d <- merge( d1, d2, by="tractid" )
d <- merge( d, md, by="tractid" )

#load NMTC data 
URL1 <- "https://raw.githubusercontent.com/DS4PS/cpp-528-spr-2020/master/labs/data/raw/NMTC/nmtc-sheet-01.csv"
nmtc <- read.csv( URL1, stringsAsFactors=F )

# remove anything not a number from the string
d$id2 <- gsub( "[^0-9]", "", d$tractid )

# fix IDs so they are match
d$id2 <- as.numeric( d$id2 )

# need to convert from currency to numeric
# current format: 
# head( nmtc$QLICI.Amount )
# [1] "$300,000.00 "   "$1,008,750.00 " "$977,000.00 "

# remove dollar sign and commas
nmtc$amount <- gsub( "[,$]", "", nmtc$QLICI.Amount )

# head(  nmtc$amount  )
# "300000.00 "  "1008750.00 " "977000.00 "

# convert characters to numeric 
nmtc$amount <- as.numeric( nmtc$amount ) %>% round(0)

# head(  nmtc$amount  )
# [1]  300000 1008750  977000

nmtc.dollars <- 
  nmtc %>% 
  filter( Origination.Year >= 2000 & Origination.Year <= 2010 ) %>%
  group_by( X2010.Census.Tract ) %>% 
  summarize( num.nmtc=n(), nmtc.total = sum( amount, na.rm=T ) )

d <- merge( d, nmtc.dollars, by.x="id2", by.y="X2010.Census.Tract", all.x=T )

# recode tracts that had no grants from NA to 0
d$num.nmtc[ is.na(d$num.nmtc) ] <- 0
d$nmtc.total[ is.na(d$nmtc.total) ] <- 0

#remove rural districts 
d <- filter( d, urban == "urban" )


```

### Create data frame for matching

```{r}
# create treatment variable
treatment <- ifelse(d$nmtc.total>0, 1, 0)

# add to dataframe

d$treatment <- treatment

```


```{r}
head(d)
```



```{r}
# match nmtc as treatment (census tracts that recieved nmtc funding vs those that didn't)
# and seek balance on other SES characteristics 

## immigrants, high school education, home ownership, pop density (cpp 528 lab), tenure in the neighborhood

df <- 
  d %>% 
  select( cbsa, id2, treatment, rent00, hinc00, pnhwht12, pnhblk12, phisp12, 
              ppov12, punemp12,
              hinc12, hincb12, hincw12, 
              phs12, pcol12,
              pown12, mhmval12 ) %>% 
 filter( treatment %in% c( 1,0 ) ) %>% 
 mutate( Treated = as.numeric( treatment == 1 ), 
           Control =  as.numeric( treatment == 0 ))

# drop rows with missing values

nrow( df )
df2 <- na.omit( df )
nrow( df2 )


# "treated" group is white
# comparison set is minority communities
```


### matchit function 

### Genetic Search Method

Better method, but more computationally intensive. 

We use a smaller sample here, but to run with the full sample it may require running it overnight or on the supercomputing clusters. 


**NOTE** the long output printed while the algorithm runs. Add code chunk option

results 

```{r, results="hide", cache=TRUE}
# test genetic search method
# is time-intensive so test with a sample

# set seeds if you want to reproduce the 
# exact result in the future
set.seed( seed=1234 )

# create sample set for cities with high NMTC funding: New Orleans
df3 <- df2  %>%  filter(cbsa %in% 35380)

row.names(df3) <- df3$id2    
df3 <- na.omit( df3 )

# df3$cbsa.f <- factor(df3$cbsa)      
                 
table( df3$treatment )


m5 <- matchit( treatment ~ hinc12 + pcol12 + pnhwht12, 
               method="genetic", discard="both", reestimate=TRUE, 
               replace=FALSE, caliper=.25, data = df3 ) 

```

Report resulting fits:

```{r}
 
summary( m5 )   # replacement false with calipers 
```


Constructing the balanced sample: 


```{r}
# match matrix is the length of the original treated groups
# IDs are rows of the original dataset

matches <- data.frame( treatment.group.id=row.names( m5$match.matrix), 
                       control.group.id=m5$match.matrix )

these.na <- is.na(matches$control.group.id)

id.all.treated <- matches$treatment.group.id 
id.treated.matched <- matches$treatment.group.id[!these.na]
id.treated.unmatched <- matches$treatment.group.id[these.na]
id.control.matched <- na.omit(matches$control.group.id)
id.control.unmatched <- df3$id2 [ !(df3$id2 %in% id.all.treated | df3$id2 %in% id.control.matched)]


matches <- na.omit(matches)

sum(is.na(matches$control.group.id))

head( matches, 61 )

```

```{r}
print(matches$treatment.group.id)
print(matches$control.group.id)
```


```{r}
# compare quality of matches
# select id (row) of treated and id (row) of comparison,
# compare on model covariates:

# note the nice balance: 

df3[ c(10,334, 14, 331, 22, 198) , c("cbsa", "id2", "hinc12", "pcol12", "pnhwht12", "treatment") ]


m5$nn

# all treated cases in dataset:
length( m5$match.matrix )

# treated cases that were matched: 
sum( ! is.na(m5$match.matrix) )
```

### Extract balanced dataset 

Keep only cases that were matched in previous steps 

```{r}
new.dat <- match.data( m5 )
dim( new.dat ) # 2 x number in matched treatment group
head( new.dat, 34 )
```

# Integrate matching code with map


### Create Treatment and Control vectors

```{r}

# id.all.treated <- matches$treatment.group.id 
# id.treated.matched <- matches$treatment.group.id[!these.na]
# id.treated.unmatched <- matches$treatment.group.id[these.na]
# id.control.matched <- na.omit(matches$control.group.id)
# id.control.unmatched <- df3$id2 [ !(df3$id2 %in% id.all.treated | df3$id2 %in% id.control.matched)]

treatment.matched <- na.omit(id.treated.matched) %>% as.numeric
control.matched <- na.omit(id.control.matched) %>% as.numeric

treated.unmatched <- na.omit(id.treated.unmatched)  %>% as.numeric
control.unmatched <- na.omit(id.control.unmatched)  %>% as.numeric

```

```{r}
head(treatment.matched)
```


```{r}
# Matched

dput(treatment.matched)
dput(control.matched)

# Unmatched 

dput(treated.unmatched)
dput(control.unmatched)
```

```{r}
# Create fips 

treatment.fips <- c(22051020502, 22051020508, 22051020600, 22051024202, 22051025900, 
22051026000, 22051027602, 22051028100, 22071000605, 22071000613, 
22071000701, 22071001302, 22071001402, 22071001500, 22071001720, 
22071001724, 22071001751, 22071001900, 22071003301, 22071003308, 
22071005000, 22071007101, 22071007200, 22071011100, 22071013600, 
22071013700, 22071014000, 22087030800, 22095070700, 22103040102, 
22103040605, 22103040803)

control.fips <- c(22089062700, 22089062302, 22051024600, 22071010600, 22051021200, 
22087030601, 22095070800, 22071000300, 22087030204, 22071002700, 
22071000901, 22071013100, 22071003500, 22051025201, 22071001746, 
22071013800, 22071009200, 22051025700, 22071012300, 22071001735, 
22071004000, 22071004900, 22071004401, 22071003701, 22071000615, 
22071003000, 22071002000, 22087030103, 22095070600, 22051024100, 
22051023903, 22051025104)

unmatched.fips <- c(22051027701, 22071000616, 22071001800, 22071002600, 22071003702, 
22071005604, 22071007000, 22071007700, 22071008400, 22071008800, 
22071009000, 22071010300, 22071010700, 22071010900, 22071011900, 
22071012102, 22071012600, 22071013200, 22071013400, 22071013900, 
22071980000, 22075050200, 22051020101, 22051020102, 22051020201, 22051020202, 22051020203, 
22051020301, 22051020302, 22051020303, 22051020400, 22051020505, 
22051020506, 22051020507, 22051020511, 22051020512, 22051020513, 
22051020514, 22051020515, 22051020516, 22051020517, 22051020700, 
22051021000, 22051021100, 22051021300, 22051021400, 22051021500, 
22051021600, 22051021700, 22051021801, 22051021803, 22051021804, 
22051021900, 22051022001, 22051022002, 22051022101, 22051022102, 
22051022200, 22051022301, 22051022302, 22051022303, 22051022400, 
22051022500, 22051022600, 22051022700, 22051022800, 22051022900, 
22051023001, 22051023002, 22051023003, 22051023100, 22051023200, 
22051023300, 22051023400, 22051023500, 22051023600, 22051023700, 
22051023800, 22051023901, 22051023902, 22051023904, 22051024001, 
22051024002, 22051024201, 22051024300, 22051024400, 22051024500, 
22051024700, 22051024800, 22051024900, 22051025001, 22051025002, 
22051025003, 22051025102, 22051025103, 22051025202, 22051025300, 
22051025400, 22051025500, 22051025600, 22051025800, 22051026100, 
22051026200, 22051026300, 22051026400, 22051026500, 22051026600, 
22051026700, 22051026800, 22051026900, 22051027000, 22051027100, 
22051027200, 22051027501, 22051027502, 22051027601, 22051027703, 
22051027803, 22051027804, 22051027805, 22051027806, 22051027807, 
22051027809, 22051027810, 22051027811, 22051027812, 22051027901, 
22051027902, 22051028000, 22051028200, 22051980000, 22071000100, 
22071000200, 22071000400, 22071000601, 22071000602, 22071000603, 
22071000604, 22071000606, 22071000607, 22071000611, 22071000612, 
22071000617, 22071000618, 22071000702, 22071000800, 22071000902, 
22071000903, 22071000904, 22071001100, 22071001200, 22071001301, 
22071001401, 22071001600, 22071001701, 22071001702, 22071001722, 
22071001723, 22071001725, 22071001730, 22071001734, 22071001736, 
22071001737, 22071001739, 22071001740, 22071001741, 22071001743, 
22071001744, 22071001745, 22071001747, 22071001748, 22071001749, 
22071001750, 22071002100, 22071002200, 22071002300, 22071002401, 
22071002402, 22071002501, 22071002502, 22071002503, 22071002504, 
22071002800, 22071002900, 22071003100, 22071003302, 22071003303, 
22071003304, 22071003307, 22071003400, 22071003600, 22071003800, 
22071003900, 22071004100, 22071004402, 22071004500, 22071004600, 
22071004800, 22071005400, 22071005500, 22071005601, 22071005602, 
22071005603, 22071006000, 22071006300, 22071006400, 22071006500, 
22071006900, 22071007501, 22071007502, 22071007604, 22071007605, 
22071007606, 22071007800, 22071008200, 22071008300, 22071008500, 
22071008600, 22071009100, 22071009400, 22071009600, 22071009700, 
22071009900, 22071010000, 22071010100, 22071010200, 22071010800, 
22071011200, 22071011400, 22071011500, 22071011600, 22071011700, 
22071012000, 22071012101, 22071012200, 22071012400, 22071012500, 
22071012700, 22071012800, 22071012900, 22071013000, 22071013301, 
22071013302, 22071013500, 22071014100, 22071014200, 22071014300, 
22071014400, 22071014500, 22071980100, 22075050100, 22075050300, 
22075050400, 22075050500, 22075050600, 22075050700, 22075050800, 
22087030104, 22087030105, 22087030203, 22087030206, 22087030207, 
22087030208, 22087030209, 22087030300, 22087030400, 22087030500, 
22087030602, 22087030603, 22087030700, 22089060100, 22089062100, 
22089062200, 22089062301, 22089062400, 22089062500, 22089062800, 
22089062900, 22089063000, 22089063100, 22089063200, 22095070100, 
22095070200, 22095070300, 22095070400, 22095070500, 22095070900, 
22095071000, 22095071100, 22103040103, 22103040104, 22103040201, 
22103040202, 22103040303, 22103040304, 22103040305, 22103040400, 
22103040501, 22103040502, 22103040601, 22103040602, 22103040604, 
22103040701, 22103040704, 22103040705, 22103040706, 22103040708, 
22103040709, 22103040710, 22103040801, 22103040802, 22103040900, 
22103041002, 22103041003, 22103041004, 22103041101, 22103041102, 
22103041103, 22103041104, 22103041202, 22103041204, 22103041207, 
22103041208, 22103041209, 22103041210, 22103041211, 22103041212, 
22103041300)
```

### Load New Orleans dorling


```{r}
library( geojsonio )
library( sp )

github.url <- "https://github.com/DS4PS/usa-dorling-shapefiles/raw/master/maps/metros-dorling/new-orleans-metairie-kenner-la-dorling-v2.geojson"
no <- geojson_read( x=github.url,  what="sp" )
 
plot( no )


```



### Create color vectors based on matches 

```{r}
color.vector <- rep( "gray", nrow(new.dat) )
color.vector[ new.dat$id2 %in% control.matched ] <- "blue"
color.vector[ new.dat$id2 %in% treatment.matched ] <- "red"

```

